---
title: "Using and Extending the Easter Egg Factory Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using and Extending the Easter Egg Factory Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

This vignette shows how to:

1. run the built-in `easter_egg_factory_v0_config()` scenario,
2. inspect node logs, and
3. extend the example with:
   - an additional supplier,
   - an additional egg recipe,
   - a new machine operator,
   - a new machine.

```{r setup}
library(codexdatasci)
```

## 1) Run the base example configuration

```{r base-config}
cfg <- easter_egg_factory_v0_config()

shape_ok <- validate_config_shape(cfg)
symbols <- validate_config_symbols(cfg)

shape_ok
symbols$ok
```

Provide required global state and node state inputs, then run.

```{r run-base}
init_global_state <- list(
  reorder_threshold = 20,
  base_rate = 6
)

init_node_state <- list(
  supplier_batches = list(n_suppliers = 3, n_ingredients = 3),
  inventory = list(start_qty = 30),
  reorder_policy = list(reorder_buffer = 10),
  operator_shifts = list(n_operators = 3),
  machine_config = list(n_machines = 3),
  staffing = list(max_ops_per_machine = 1),
  production = list(quality_floor = 0.75)
)

logs <- run_simulation(
  config = cfg,
  init_global_state = init_global_state,
  init_node_state = init_node_state
)

names(logs)
```

Inspect a few rows from the machine and production logs.

```{r inspect-base}
head(logs$machine_config, 6)
head(logs$production, 6)
```

## 2) Extend the scenario

To extend the scenario, we update both state counts and the relevant node
expressions.

### 2.1 Add one supplier and one operator

We will move from 3 to 4 suppliers/operators by changing node state values.

### 2.2 Add one egg recipe and one machine

We will move from 3 to 4 machines and add a fourth recipe:

- `ruby_caramel` using `ING1+ING4`.

Because this recipe uses a new ingredient (`ING4`), we also increase the
ingredient pool from 3 to 4.

```{r extend-config}
ext_cfg <- easter_egg_factory_v0_config()

# Add a 4th supplier profile (SUP4) for price and quality behavior.
ext_cfg$nodes$supplier_batches$expressions$batch_price <-
  "ifelse(supplier_id == 'SUP1', 1.2, ifelse(supplier_id == 'SUP2', 1.1, ifelse(supplier_id == 'SUP3', 1.0, 0.98)))"
ext_cfg$nodes$supplier_batches$expressions$batch_quality <-
  "ifelse(supplier_id == 'SUP1', 0.95, ifelse(supplier_id == 'SUP2', 0.90, ifelse(supplier_id == 'SUP3', 0.86, 0.91))) + runif(1, -0.01, 0.01)"

# Add M4 and a new recipe (ruby_caramel).
ext_cfg$nodes$machine_config$expressions$egg_type <-
  "ifelse(machine_id == 'M1', 'dark_truffle', ifelse(machine_id == 'M2', 'milk_hazelnut', ifelse(machine_id == 'M3', 'white_berry', 'ruby_caramel')))"
ext_cfg$nodes$machine_config$expressions$machine_quality <-
  "ifelse(machine_id == 'M1', 0.93, ifelse(machine_id == 'M2', 0.89, ifelse(machine_id == 'M3', 0.87, 0.91)))"
ext_cfg$nodes$machine_config$expressions$required_ingredients <-
  "ifelse(egg_type == 'dark_truffle', 'ING1+ING2', ifelse(egg_type == 'milk_hazelnut', 'ING2+ING3', ifelse(egg_type == 'white_berry', 'ING1+ING3', 'ING1+ING4')))"

# Add OP4 quality profile.
ext_cfg$nodes$operator_shifts$expressions$operator_quality <-
  "ifelse(operator_id == 'OP1', 0.92, ifelse(operator_id == 'OP2', 0.88, ifelse(operator_id == 'OP3', 0.84, 0.90)))"

ext_symbols <- validate_config_symbols(ext_cfg)
ext_symbols$ok
```

Now run the extended configuration with updated node state counts.

```{r run-extended}
ext_node_state <- list(
  supplier_batches = list(n_suppliers = 4, n_ingredients = 4),
  inventory = list(start_qty = 30),
  reorder_policy = list(reorder_buffer = 10),
  operator_shifts = list(n_operators = 4),
  machine_config = list(n_machines = 4),
  staffing = list(max_ops_per_machine = 1),
  production = list(quality_floor = 0.75)
)

ext_logs <- run_simulation(
  config = ext_cfg,
  init_global_state = init_global_state,
  init_node_state = ext_node_state
)
```

## 3) Verify the new entities appear

```{r verify-extensions}
unique(ext_logs$supplier_batches$supplier_id)
unique(ext_logs$operator_shifts$operator_id)
unique(ext_logs$machine_config$machine_id)
unique(ext_logs$machine_config$egg_type)
unique(ext_logs$machine_config$required_ingredients)
```

You should see `SUP4`, `OP4`, `M4`, and the new recipe `ruby_caramel` with
`ING1+ING4` in the output.
